Teaching Dossier

### 2020-02-22 CUT/teaching-philosophy.tex
Future goals

\paragraph{}
\subsection{Assessment}

\paragraph{} \textit{Importance of spaced repetition for studying. Importance of priming students for lecture via problems. Assigment problems and timing}
\begin{itemize}
    \item Multi-modal assessments
    \begin{itemize}
        \item Classic problem sets
        \item Oral interviews
        \item Long term projects with group collaboration
        \item Mdterms and Final Exam with cheatsheet format.
    \end{itemize}
    \item Oral vs Written exams
    \item Importance of Collaboration and long term projects
    \item The cheatsheet format for exams and its benefits
\end{itemize}
\section{Evaluation of Teaching}
I find feedback on my teaching to be tremendously valuable. It is the easiest way for me to identify potential avenues for improvement, while at the same time motivating me by highlighting strong points of my teaching style. I have organized my evaluations below into two categories: those from students in courses I have taught, and those from peers and teaching professionals observing my lectures.
\subsection{Student Evaluations}
%Chart out survey responses for CO 327 with survey adjectives for context for scores (some here, rest in appendix)
%Student comment strong points
%Student comment targets for change
\subsection{Teaching Observations}
One element of the CUT teaching program I took alongside my PhD was having lectures I give upervised by an expert in teaching working with the university's Centre for Teaching Excellence (CTE). I completed two teaching observations through the program. Below I summarize the feedback I received from these observations.
%highlights from 1st observation
\subsubsection*{First Teaching Observation}
\paragraph{}
My first observation was on the guest lecture I gave for Introduction to Game Theory in Fall $2018$ on Correlated Equilibria. The observation was conducted by Shahrukh Athar, a Graduate Instructional Developer working for the CTE. I have included the relevant sections of the full report in Appendix \ref{sec:observation-report-1}. The feedback in the report was divided in \emph{aspects to maintain}, where strengths of teaching were identified, and \emph{targets for change}, where opportunities for improvement were identified.
\paragraph{Aspects to Maintain} The first strength Shahrukh identified was my use of questions and comprehension checks. He appreciated the variety of checks I used, as well as the split between content-specific checks and general purpose open-ended questions. He commented that ``\emph{All of this is
fantastic as it demonstrates a learner-centric approach to teaching rather than a teacher-centric
one. It shows that you want to take your learners along with you as the lesson progresses and
are willing to pause and clarify confusions before moving on.}''

@@

Caelan's general impression of the lesson can be summarized as ``\emph{You were confident, clear, and
knowledgeable in the classroom. Overall, your lesson was definitely a success}.''. As with the previous report, more detailed feedback was separated into aspects to maintain and targets for change. The relevant sections of the full report can be found in Appendix \ref{sec:observation-report-2}
\paragraph{Aspects to Maintain}

\paragraph{Targets for Change}
%highlights from 2nd observation with a narrative about improving

\section{Professional Development}
%FUT
%Math Graduate Teaching Seminar
%CUT
\section{Future Goals}
%Write a textbook
%Teach a fourth year course
%Teach a topics course
%Become more invovled in teaching 1st and 2nd years
%Head instructor for a course with multiple sections
%Graduate and URA supervision

\appendix
\section*{Appendices}
%Full evaluations
%Observation Reports
%Sample Lesson Plan from guest lecture
%Sample Assignment from CO 327
%Sample Lesson Plan from CO 327
\section{Teaching Observation Report 1}\label{sec:observation-report-1}

>>>>

\paragraph{}
\subsection{Assessment}
\paragraph{}
My approach to assessment is based not only only my experience as a teacher but on my understanding of the research literature. During the CUT program, I wrote a research paper surveying the academic literature on how exam formats impact student learning outcomes. Some interesting findings were synthesized from this paper which I now incorporate into my approach to student assessment.
\paragraph{}
Written midterms and final exams are fundamental evaluations which occur during courses I teach. They are usually cummulative and are worth a significant portion of a student's final grade. Naturally this means that many students will experience so-called test anxiety while preparing for and taking these exams. The literature show that offering students open book exams, i.e.\ exams where they are allowed to reference textbooks and/or course notes, significantly reduces test anxiety. Unfortunately the literature also shows that open book exams lead to significantly lower long term retention of course material than closed book exams. Interestingly, there is a way to obtain the same test anxiety reducing benefits of an open book exam while maintaining the long term retention of a closed book exam. This way is the cheatsheet exam, which I use in all my courses. For exams which do not have a predetermined format decided by institutional policy I allow students to bring into the exam a small cue card on which they can write whatever they want. The literature shows that this format has comparable reductins in test anxiety among students as an open book exam, while also encouraging postive study habits which lead to long term retention of the material since students are forced to prioritize the information they write on their cue card.
\paragraph{}
Alongside written exams I like to encorporate an oral exam component into my courses. This is implemented in an interview format, where a student is given a topic present on, such as a significant theorem from the course, they then prepare their presentation and schedule with me a time where they give the presentation and I ask questions to probe the depths of their understanding. I like to give students a list of topics from which they can engage in self-directed study for this oral exam. The research literature shows that oral exams not only improve student confidence, encourage effective study habits, and strengthen communication skills, but that performance on oral exams is only weakly correlated with performance on written exams. This suggests that they test a different set of skills. In the interest of celebrating the strengths of a diverse student body I strive to incorporate exam formats like oral exams which give students who may struggle with written exams an opportunity to demonstrate their knowledge and learning attainment.
\paragraph{}
Whatever plans students may have after graduation they will likely need to be able to work as part of a team. Increasingly mathematics research is done collaboratively, and in industry every professional is embedded as part of a collaborative team. Thus I believe it is important for students to have opportunities to work together in their courses. I usually give students this opportunity in the form of a long term open ended group project. For instance, in a course on optimization students may be given a more applied project based on an industry case study. Each group would be tasked with proposing a model for the problem and solving it using techniques from the course in whatever way they see fit. This encourages them to not only develop their ability to work with others, but also to leverage their creativity. Finally, it gives them a chance to see how the mathematics they learn in the classroom can have real world impact.

\section{Evaluation of Teaching}
I find feedback on my teaching to be tremendously valuable. It is the easiest way for me to identify potential avenues for improvement, while at the same time motivating me by highlighting strong points of my teaching style. I have organized my evaluations below into two categories: those from students in courses I have taught, and those from peers and teaching professionals observing my lectures.
\subsection{Student Evaluations}
%Chart out survey responses for CO 327 with survey adjectives for context for scores (some here, rest in appendix)
%Student comment strong points
%Student comment targets for change
\subsection{Teaching Observations}\label{sec:teaching-observations}
One element of the CUT teaching program I took alongside my PhD was having lectures I give upervised by an expert in teaching working with the university's Centre for Teaching Excellence (CTE). I completed two teaching observations through the program. Below I summarize the feedback I received from these observations.
%highlights from 1st observation
\subsubsection*{First Teaching Observation}
\paragraph{}
My first observation was on the guest lecture I gave for Introduction to Game Theory in Fall $2018$ on Correlated Equilibria. The observation was conducted by Shahrukh Athar, a Graduate Instructional Developer working for the CTE. I have included the relevant sections of the full report in Appendix \ref{sec:observation-report-1}. The feedback in the report was divided in \emph{aspects to maintain}, where strengths of teaching were identified, and \emph{targets for change}, where opportunities for improvement were identified.
\paragraph{Aspects to Maintain} The first strength Shahrukh identified was my use of questions and comprehension checks. He appreciated the variety of checks I used, as well as the split between content-specific checks and general purpose open-ended questions. He commented that ``\emph{All of this is
fantastic as it demonstrates a learner-centric approach to teaching rather than a teacher-centric
one. It shows that you want to take your learners along with you as the lesson progresses and
are willing to pause and clarify confusions before moving on.}''

@@

Caelan's general impression of the lesson can be summarized as ``\emph{You were confident, clear, and
knowledgeable in the classroom. Overall, your lesson was definitely a success}.''. As with the previous report, more detailed feedback was separated into aspects to maintain and targets for change. The relevant sections of the full report can be found in Appendix \ref{sec:observation-report-2}
\paragraph{Aspects to Maintain}
Caelan appreciated the structure of my lesson and my preparation. Specifically she highlighted my use of a lecture outline and my starting the session with a review of the previous lecture's material connecting that content to the upcoming lecture. She also appreciated my confident delivery, saying ``\emph{in fact, you did not stumble even once in your teaching, which was impressive.}''

Caelan was also impressed with my handwriting and board use. She found the presentation very clear and liked the use of emphasis with bullet points and underlining, as well as the strategic positioning of figures so that they can stay on the board as long as possible given the small amount of board space in the particular room the lecture took place in.

Finally, Caelan celebrated my use of examples and questioning strategies. She liked how I presentated examples to motivate upcoming theory and definitions, and would revisit examples throughout the lecture to help reinforce new ideas as the lecture progressed. She said ``\emph{Such well-thought examples and questions really kept your learners actively engaged during your lesson. I would highly encourage you to continue these
great practices in your future teaching!}''.
\paragraph{Targets for Change}
%highlights from 2nd observation with a narrative about improving
One suggestion for improvement Caelan had was to remember to repeat questions or responses given by students in class. Her argument being that even though the classroom was small students are concentrating on the instructor, not necessarily each other, and it is helpful to do this to make sure they hear important information. I agree with her assessment, and this is an easy thing to tweak in my lecturing technique.

I had asked Caelan to pay attention to the Write-Turn-Talk technique I was working on from Shahrukh's feedback on my previous observation. She noted that I had successfully implemented the habit of not talking while writing on the board, but she had a related taret for change. While I do not write and talk, I would sometimes face the board to point out aspects of a definition or elements of a figure I had drawn in relation to a student's question. She suggested an adjustment where I practice facing the audience while pointing to elements of the figure. This can be tricky, but the idea would be to look at the figure, point out what I have in mind, turn and then discuss it. In future lectures I plan to practice this technique and I think it will certainly help strengthen communication with my students.

\section{Professional Development}
\subsection{Fundamentals of University Teaching}
\paragraph{}
During my Masters degree I completed the Fundamentals of University Teaching program from the CTE at the University of Waterloo. The program consisted of six workshops which focused effective lesson plan creation, questioning strategies, designing student assessments, motivating students, managing student atteiont, and giving and receiving feedback. Also as part of the program I gave three micro-teaching sessions to my peers in the program. A microteaching session consists of brief $15$ minute lecture with an emphasis on active learning techniques and student engagement. Each session was evaluated by an affiliate of the CTE from whom I received feedback that identified aspects of my teaching to maintain and targets for change. This program provided an excellent foundation for my approach to university teaching.

\subsection{Certificate in University Teaching}
\paragraph{}During my PhD studies I had an opportunity to expand upon the teaching skills I developed in the Fundamentals of University Teaching via the Certificate in University Teaching program. This program is more intensive and in-depth than the Fundamental program. This program involved four workshops: understanding the learner, interactive teaching strategies, assessing student learning, and course design. Each workshop required me to complete a response paper where I reflected on how the concepts from the workshop could apply in my teaching. The program also involved two in-class teaching observations by a CTE Graduate Instructional Developer, the resulting feedback from which I discuss in Subsection \ref{sec:teaching-observations}.

For the program I wrote a research paper summarizing the current research literature on ``Impact of Exam Format on Student Learning Outcomes''. An extended abstract can be found in the Appendices \textbf{TODO LINK}.

\subsection{Mathematics Teaching Seminar} 
\paragraph{}
The Faculty of Mathematics at the University of Waterloo offers a $12$-week program for graduate students to develop their teaching skills. Each week involved a $2$ hour seminar on an aspect of teaching skills with a specialized focus on teaching mathematics. I completed this program as part of my PhD.

Some specialized seimar presentations studied topics such as presentations skills with a particular focus on blackboard technique for mathematics,  and designing mathematics problem sets. The program also involved microteaching sessions similar to the Fundamentals of University Teaching Program, but since the peers were all graduate students in mathematics the lectures could be tuned to a more specialized audience to practice more technical skills. It also invovled teaching observations of some of the most highly reviewed lectures at Waterloo, and reflection discussions on what techniques they are using to be effective, how they engage students, and how they deal with setbacks during the lecture. Overall this seminar series was very valuable in supplementing the skills I had developed through the CTE programs with techniques specifically focused on the discipline in which I will be teaching.

\section{Future Goals}

\paragraph{}
My immediate short term goal is to teach Deterministic OR Models again in Spring $2020$. I had a great experience last time I taught it, and the students had very positive feedback, but there some suggestions for improvement. I'm hoping to incorporate the feedback I have received to improve the experience for students during the next offering.

Thinking more long term I would like to become a professor at an institution which values high quality teaching as much as research. I am excited to teach across all levels of undergraduate and graduate students. Teaching students early in their journey is a great opportunity to give them a solid foundation in the fundamentals they will need to matter what direction they choose and to get them excited and passionate about pursuing mathematics further. Teaching more senior students and graduate students is a great opportunity to have more close interaction with each individual student, to share my specialities in mathematics, and to expose students to cutting edge research topics.

I would also like to participate more in outreach to high school students, and possibly even younger students. I believe my areas of focus in algorithms, combinatorics, and optimization have many topics which are highly accessible to students who have not yet seen calculus. These topics are full of many interesting puzzles and are also imminently applicable across a wide range of industries, in other words they are perfect for engaging young people and encouraging them to pursue mathematics.

\appendix
\section*{Appendices}
%Observation Reports
%Extended Abstract for Research paper
%Full evaluations
%Sample Lesson Plan from guest lecture
%Sample Assignment from CO 327
%Sample Lesson Plan from CO 327
\section{Teaching Observation Report 1}\label{sec:observation-report-1}
